---
title: "Fitting a single species Poisson process model."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting a single species Poisson process model.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ppmData)
library(ppmFit)
library(glmnet)
```


Firstly we need to set up a ppmData object. This can be done using the ppmData function.

```{r run ppmData, cache=TRUE}
path <- system.file("extdata", package = "ppmData")
lst <- list.files(path=path,pattern='*.tif',full.names = TRUE)
covariates <- rast(lst)
presences <- subset(snails,SpeciesID %in% "Tasmaphena sinclairi")
ppmdata <- ppmData(npoints = 10000,
                   presences = presences,
                   window = covariates[[1]],
                   covariates = covariates,
                   na.rm = TRUE)
```

```{r plot ppmdat}
plot(ppmdata)
```

```{r fit model}
sp_form <- presence ~ poly(X,2) + poly(Y,2) + poly(max_temp_hottest_month,2)+
           poly(annual_mean_precip,2) + poly(annual_mean_temp,2) +
           poly(distance_from_main_roads,2)

ft.ppm <- ppmFit(species_formula = sp_form,
                 ppmdata=ppmdata,
                 method='lasso',
                 family = "poisson")
```

Run the cross validation from glmnet to get the best value for lambda. Lambda in the glmnet sense, is the value of the penalty in the elastic net algorithm.
```{r cv lambda}
cvl <- cvLambda(ft.ppm)
```

Plot the lambda cross-validation if so desired. The two lines identify the "lamdba.min" which gives the value of lambda that gives minimum cross-validated error and "lambda.1se" which gives largest value of lambda such that error is within 1 standard error of the minimum. I typically use the "lambda.min" value for prediction, but any choice of lambda is possible.
```{r cv lambda plot}
plot(cvl)
```

Predict the model to the covariates 
```{r}
pred <- predict(ft.ppm,newdata=covariates)
plot(pred)
```

